Model validation plots

### 2021-08-27 doc/validate_models.Rmd
finish model validation supplement

**TODO**

- check that # of knots are adequate
- check diagnostics for individual-level random effects.  Doesn't look good, but not sure how to interpret or what to do.

@@

Basis dimension checking with `gam.check()` doesn't appear to work for `dlnm` crossbasis smooths. Instead I'll use a method described in the help file `?choose.k` to check for adequate knots.  Unfortunately, `bs = "cs"` doesn't work with `dlnm`, so I'll use `select = TRUE` instead to reduce chance of overfitting.

```{r eval=TRUE}
check_res_edf(s_1ha)
```

The crossbasis smooth possibly needs more knots.

```{r}
# plot_lag_slice(s_1ha, "spei_history", lag =  0) + labs(title = "surv, 1ha")
```

Here you can see how the CIs hardly overlap 0, but it could be an artifact of the line not being allowed to be wiggly enough.

### Summary

```{r eval=TRUE}
summary(s_1ha)
```

```{r eval=TRUE}
draw(s_1ha)
```

@@

```{r}
# looking for near zero edf
check_res_edf(s_cf)
```

### Summary

```{r}
summary(s_cf)
```

```{r}
draw(s_cf)
```

@@

Here I use a random sub-sample to check that differences between continuous forest and fragments are not purely due to sample size differences, particularly differences in the complexity of the crossbasis smooth.

```{r}
summary(s_1ha)$edf[3]
summary(s_cf)$edf[3]
summary(s_cf_sub)$edf[3]
```

```{r}
draw(s_cf_sub)
```
Surface still looks different in a similar way though.

@@

@@

### Summary

```{r}
summary(g_1ha)
```

```{r}
draw(g_1ha)
```

@@

@@

```{r}
# plot_lag_slice(g_1ha, "spei_history", lag = 33)
```

hmm... CIs are **super** narrow here, but effect size is **tiny**

### Summary

```{r}
summary(g_cf)
```

```{r}
draw(g_cf)
```

## Continuous Forest subsample

@@

```{r}
summary(g_1ha)$edf[3]
summary(g_cf)$edf[3]
summary(g_cf_sub)$edf[3]
```
edf of subsample is similar to full dataset, despite differences in sample size.

```{r}
draw(g_cf_sub)
```
Crossbasis surface looks nearly identical.

@@

### Summary

```{r}
summary(f_1ha)
```

```{r}
draw(f_1ha)
```

@@

```{r}
# looking for near zero edf
check_res_edf(f_cf)
```

### Summary

```{r}
summary(f_cf)
```

```{r}
draw(f_cf)
```

## Continuous Forest subsample

To check that differences are not purely due to sample size differences, particularly that lower edf in continuous forests is due to higher sample size.  For flowering, edf is actually slightly higher in CF with a larger sample size.  With subsample, edf are more similar.

```{r}
summary(f_1ha)$edf[3]
summary(f_cf)$edf[3]
summary(f_cf_sub)$edf[3]
```

```{r}
draw(f_cf_sub)
```
Crossbasis surface looks extremely similar.

>>>>

@@

`k` is adequate for `log_size_prev`. Unfortunately `k.check()` doesn't work for data with matrices, so I'll use an alternative approach detailed in `?mgcv::choose.k` for the crossbasis smooth.  Briefly, it involves looking for pattern in residuals by extracting residuals and fitting a GAM to them with the smooth of interest with an increased value for `k`.  If there is no unexplained pattern in residuals, that smooth should have an edf near zero.

```{r eval=TRUE}
check_res_edf(s_1ha)
```
Adequate k for crossbasis.

@@

Adequate knots for `log_size_prev`

```{r}
check_res_edf(s_cf)
```
Adequate knots for crossbasis smooth.

@@

Here I use a random sub-sample to check that differences between continuous forest and fragments are not purely due to sample size differences, particularly differences in the complexity and shape of the crossbasis smooth.

```{r}
summary(s_1ha)$edf[2]
summary(s_cf)$edf[2]
summary(s_cf_sub)$edf[2]
```

The crossbasis is much smoother

```{r}
draw(s_cf, select = "s(spei_history,L)", n_contour = 5)
draw(s_cf_sub, select = "s(spei_history,L)", n_contour = 5)
```

@@

Adequate k for `log_size_prev`.

@@

Adequate k for crossbasis.

@@

Adequate k for `log_size_prev`

@@

Adequate k for crossbasis smooth.

@@

```{r}
summary(g_1ha)$edf[2]
summary(g_cf)$edf[2]
summary(g_cf_sub)$edf[2]
```

edf is similar for the sub-sample

```{r}
draw(g_cf, select = "s(spei_history,L)", n_contour = 5)
draw(g_cf_sub, select = "s(spei_history,L)", n_contour = 5)
```

The shape of the smooth is also similar.

@@

Adequate k for crossbasis smooth.

@@

Adequate k for `log_size_prev`

```{r}
# looking for near zero edf
check_res_edf(f_cf)
```
Adequate k for crossbasis.

```{r}
summary(f_1ha)$edf[2]
summary(f_cf)$edf[2]
summary(f_cf_sub)$edf[2]
```

Similar edf for sub-sample

```{r}
draw(f_cf, select = "s(spei_history,L)", n_contour = 5)
draw(f_cf_sub, select = "s(spei_history,L)", n_contour = 5)
```

Similar, but essentially linear in SPEI dimension.
